Interpretable Neural Networks using EAGGA

### 2025-03-15 paper/main.tex
MOO pareto front + a priori

\subsection{Multi-Objective Optimization}
- in most practical use cases one doesn't just want to optimise for performance but e.g. also for interpretability using some proxy metrics
- formal definition
  * vector of objectives $c_1,c_2,...,c_m=\boldsymbol{c}:\boldsymbol\Lambda\rightarrow\mathbb{R}^m$
  * goal: minimise vector
  * \citep[p. 11]{10.1145/3610536}
\subsubsection{Pareto-optimality}
- problem: usually conflicting objectives, i.e. minimisation of $\boldsymbol{c}$ not possible along all dimensions
- thus aim to find trade-off solutions of non-dominated points
- we say a ``point dominates another''
  * if there is no other point strictly better in at least one dimension and better or equal in the remaining ones
  * formally: $\boldsymbol\lambda$ dominates $\boldsymbol\lambda'$ ($\boldsymbol\lambda\prec\boldsymbol\lambda'$) if and only if
    $\forall i\in\{1,...,m\}:c_i(\boldsymbol\lambda) \le c_i(\boldsymbol\lambda') \wedge \exists j\in\{1,...,m\}:c_j(\boldsymbol\lambda) < c_j(\boldsymbol\lambda')$
  * \citep[pp. 7f]{10.1145/3610536} + \citep[pp. 198f]{genetic_algos}
- Pareto set: set of nondominated points $\mathcal{P}:=\{\boldsymbol\lambda\in\boldsymbol\Lambda|\nexists\boldsymbol\lambda'\in\boldsymbol\Lambda\text{ s.t. }\boldsymbol\lambda'\prec\boldsymbol\lambda\}$
- Pareto front: image of nondominated points
- goal: find set of nondominated points $\hat{\mathcal{P}}$ approximating true Pareto set $\mathcal{P}$ well
(evaluation)
- if knowledge over true Pareto-front, evaluating set of individuals can be done based on distance of estimated to true Pareto front
- if no knowledge over true Pareto-front, volume-based approaches are popular, which measure volume between Pareto front estim and some chosen
  reference point (usually worst point in objective space), e.g. 0 for accuracy (optimum is 1)
- \citep[pp. 8-10]{10.1145/3610536}
\subsubsection{A-priori}
- requires specifying trade-off between objectives a-priori, will outline 2 popular approaches
- e.g. different versions of \textbf{scalarization}
  * weighted sum of objective functions
    - $\argmin_{\boldsymbol\lambda\in\boldsymbol\Lambda} w_i c_i(\boldsymbol\lambda)$ with $\sum_{i=1}^k w_i=1$ and $w_i>0,\forall i=1,...,k$
    - drawbacks
      * solution sensitive to weights
      * different users might have different opinions on weights \citep[chap. 3.1]{NSGA}
    - \citep[p. 11]{10.1145/3610536}
  * $\epsilon$-constraint
    - translate all but one objective into constraint, then optimise remaining objective subject to the constraints
    - $\text{w.l.o.g.} \argmin_{\boldsymbol\lambda\in\boldsymbol\Lambda} c_1(\boldsymbol\lambda) \text{ s.t. } c_2(\boldsymbol\lambda)\le\epsilon_2,...,c_m(\boldsymbol\lambda)\le\epsilon_m$
    - conceputally similar to weighted sum: also sensitive to constraints, must be chosen sensibly
    - \citep[p. 12]{10.1145/3610536}
- \textbf{lexicographic method}
  * define priority of objectives
  * greedily optimise objectives in order of priority, constraint to the solutions of the already optimised (higher-priority) objectives
  * again very dependent on user-defined priorisation
  * \citep[p. 13749]{lexicographic_MOO}  % TODO: perhaps find a different source more related to ML

>>>>

\subsection{Multi-Objective Optimization}
In most applications, practicioners don't just want to optimize for performance exclusively but, for instance, also desire an interpretable model,
which they may measure via some proxy metric.
Let $c_1,c_2,...,c_m = \boldsymbol{c}:\boldsymbol\Lambda\rightarrow\mathbb{R}^m$ be the vector of $m$ objectives, the goal of multi-objective optimization (MOO)
is to minimise this vector \citep[p. 11]{10.1145/3610536}.

\subsubsection{Pareto-optimality}
Optimizing $\boldsymbol{c}$ usually comes with the challenge of conflicting objectives: improving one objective means deteriorating another.
We thus seek to find a set of trade-off solutions, so called \textit{non-dominated} points.
A point $\boldsymbol\lambda$ dominates another point $\boldsymbol\lambda'$ ($\boldsymbol\lambda\prec\boldsymbol\lambda'$) if there is no other point that is
strictly better in at least one objective and better or equal in the remaining ones. Formally,
\begin{equation}
  \forall i\in\{1,...,m\}:c_i(\boldsymbol\lambda) \le c_i(\boldsymbol\lambda') \wedge \exists j\in\{1,...,m\}:c_j(\boldsymbol\lambda) < c_j(\boldsymbol\lambda')
  \label{eq-pareto-domination}
\end{equation}
as defined in \citep[pp. 7f]{10.1145/3610536} and \citep[pp. 198f]{genetic_algos}.
\\
The set of non-dominated or \textit{Pareto-optimal}
points $\mathcal{P}:=\{\boldsymbol\lambda\in\boldsymbol\Lambda|\nexists\boldsymbol\lambda'\in\boldsymbol\Lambda\text{ s.t. }\boldsymbol\lambda'\prec\boldsymbol\lambda\}$
is commonly referred to as the \textit{Pareto set} and its image as the \textit{Pareto front}.
The goal of MOO is to find a set of non-dominated points $\hat{\mathcal{P}}$ approximating the true Pareto set $\mathcal{P}$ well.

% evaluation
There are different techniques to evaluate an estimated Pareto-front. If we have knowledge over the true Pareto-front, we can compute the distance between the approximation
and the true front. In most cases, this knowledge cannot be assumed.
Then, usually volume-based approaches are used for evaluation, most notably the hypervolume. The hypervolume, or S-metric, of a Pareto-front, computes the volume between
the estimated front and some chosen reference point, which is usually the worst point in objective space.
The larger the hypervolume, the closer the estimate is to the true Pareto-front. \citep[pp. 8-10]{10.1145/3610536}

\subsubsection{A-priori}
For optimizing a multi-objective problem, there are largely two general approaches.
One of those are \textit{a-priori} methods, of which we will briefly outline two popular techniques.

\textbf{Scalarization} approaches impose an implicit order of priority on the objectives.
The simplest one is optimizing a weighted sum of objective functions
\begin{equation}
  \argmin_{\boldsymbol\lambda\in\boldsymbol\Lambda} \sum_{i=1}^m w_i c_i(\boldsymbol\lambda)
  \text{ s.t. } \sum_{i=1}^m w_i=1 \wedge w_i>0,\forall i=1,...,m
  \label{eq-a-priori-scalarization-weighted-sum}
\end{equation}
Weights are chosen a-priori by the user and the solution is very sensitive to them, making this method very reliant on different
users' preferences. \citep[p. 11]{10.1145/3610536} and \citep[chap. 3.1]{NSGA}
\\
Another form of scalarization is the $\epsilon$-constraint, which translates all but one objective into constraints and then optimizes the remaining
objective subject to the constraints. Without loss of generality, the first constraint can be optimized
\begin{equation}
  \argmin_{\boldsymbol\lambda\in\boldsymbol\Lambda} c_1(\boldsymbol\lambda)
  \text{ s.t. } c_2(\boldsymbol\lambda)\le\epsilon_2,...,c_m(\boldsymbol\lambda)\le\epsilon_m
  \label{eq-a-priori-scalarization-epsilon}
\end{equation}
This method is, similarly to the weighted sum, conceptually simple yet also very sensitive to the chosen constraints. \citep[p. 12]{10.1145/3610536}
\\
Alternative to scalarization is the \textbf{lexicographic method}.
For it, the user assigns a priority to each objective and then greedily optimizes each objective in order of priority, constrained to the solutions of
the already optimized, higher-priority objectives. \citep[p. 13749]{lexicographic_MOO}
Again, solutions are very dependent on the user-defined priorisation.
